\propertitle

### 2019-12-15 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-15T00:52:50+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Step~\ref{item:sample_step} is just an application of the probability-calculus, which gives an exact linear relation between the first $m$
moments for the larger population and the first $m$ for the sample
\citep[Appendix]{portamanaetal2015}. The ones determine the others and
vice versa at every time bin. This relation holds for any belief
distribution $\pxxx(\yA_{t})$ for the larger-population activity and its
marginal $\pxx(\ya_{t})$ for the sample activity at that bin. These two
distributions are related by
\begin{equation}
  \label{eq:conditional_hypergeometric_t}
  \pxx(\ya_{t})= \sum_{\yA_{t}=0}^{N} \yG_{\ya_{t}\,\yA_{t}} \;\pxxx(\yA_{t})
  \qquad \ya_{t} \in \set{0,\dotsc,n},
\end{equation}
where $\yG_{\ya\yA}$ is the hypergeometric distribution defined in
\eqn~\eqref{eq:conditional_hypergeometric}, characteristic of
\enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}.%[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.

@@

In step~\ref{item:maxent_step} we actually use the
\emph{minimum-relative-entropy} method
\citep{hobsonetal1973,csiszar1985}[\sect~5.2.2]{sivia1996_r2006} with
respect to a uniform reference distribution. We shall still call it
\enquote{maximum-entropy} for brevity. It amounts to two prescriptions:
first, take the distributions satisfying specific convex constraints --
such as fixed expectations -- and among them select the one having minimum
relative entropy with respect to a reference distribution; second, judge --
and therefore set -- those expectations to be equal to some measured
averages, typically time averages. (A judgement is required because
probability expectations and empirical averages are numerically
different in general; a misjudgement of their approximate equality can lead to
\enquote{silly answers}, to quote MacKay \citep[p.~308]{mackay1995_r2003}.)

@@

The amount and degrees of the constraining moments depend on the questions
and hypotheses that a researcher is exploring. We give some examples in the
next two sections. Note that the constraint
equation~\eqref{eq:final_constraint} for the
distribution~\eqref{eq:app_maxent_pop_first} may not have exact solutions
if $N$ is strictly larger than $n$. The possible discrepancy between its
two sides typically increases with the number of constraints. This
discrepancy comes from an approximation implicit in the maximum-entropy
method combined with one assumption specific to our application: that
measured averages equal expected values (or equivalently, the number of
time bins $T$ is infinite), and that an activity level $\yA$ can equally
likely be generated by any set of $\yA$ neurons in the larger population.
The magnitude of this discrepancy can be a signature of the presence of
neuronal \enquote{assemblies} or \enquote{clusters}
\citep[\chap~12]{gerstneretal2014}{hebb1949_r2002}. We discuss this matter
in \sect~\ref{sec:assumptions}.

@@

The first assumption, of infinite exchangeability
\citep[\chap~4]{bernardoetal1994_r2000}{dawid2013} of the probability
distribution $\pf[(\yA_{1},\dotsc, \yA_{T}) \| \yI]$ for the sequence of
activities, implies the following mixture form by de~Finetti's theorem
\citep{definetti1930,hewittetal1955}:
\begin{equation}
  \label{eq:mixture_from_definetti}
%  \exp[-T\;\sh(\yF;\ynu)] \quad
%  \pxxx[(\yA_{1},\dotsc,\yA_{T})] \propto
  \pf[(\yA_{1},\dotsc, \yA_{T}) \| \yI]
  = \int\!\!\!\di\ynu\;\pf(\ynu \|\yI)\,\prod_{\yA=1}^{N}{\nu_{\yA}}^{T\,\yFF_{\yA}},
\end{equation}
where $\ynu\defd(\nu_{\yA})$ is the long-run frequency distribution of
activities in an imaginary continuation or repetition of the recording, and
$\pf(\ynu \|\yI)\;\di\ynu$ our belief distribution about it.

The second assumption concerns the latter degree distribution. It has one
of these expressions:
\begin{subequations}\label{eq:maxent_prior}
  \begin{equation}
    \label{eq:entropic_prior}
    \pf(\ynu \| \yI) \propto  \exp[-\yL\;\sh(\yF;\yr)]
    \approx \binom{L}{L\nu_{0}, \dotsc, L\nu_{N}}\;
    \prod_{\yA}{r_{\yA}}^{L\nu_{\yA}}  
  \end{equation}
  where $\yr\defd(r_{\yA})$ is a reference distribution and the expression
  in larger parentheses is a multinomial coefficient, or
  \begin{equation}
    \label{eq:suff_prior}
    \pf(\ynu \| \yI) \propto
    \int\!\!\!\di\yla\;\pf(\yla \| \yI)\;
    \prod_{\yA}\delt\Biggl\{\ynu_{\yA}-
    \frac{r_{\yA}}{Z(\yla)}\,\exp\Bigl[
    \tsum_{m}\lambda_{m}\tbinom{\yA}{m}/\tbinom{N}{m}
    \Bigr]\Biggr\},
  \end{equation}
\end{subequations}
where $\yla\defd(\lambda_{m})$ are Lagrange multipliers and $Z(\yla)$ a
normalization factor for the exponential distribution within the delta.
This is a formal expression saying that our belief about $\ynu$ is an
exponential parametric model for which the normalized factorial moments are
informationally sufficient.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-15T16:26:55+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Step~\ref{item:sample_step} is just an application of the
probability-calculus, which gives an exact linear relation between the
first $m$ moments for the larger population and the first $m$ for the
sample \citep[see][for a complete discussion and proofs. Some proofs are
summarized again here in
appendix~\ref{sec:derivation_maxent_fullnet}]{portamanaetal2015}. The ones
determine the others and vice versa at every time bin. This relation holds
for any belief distribution $\pxxx(\yA_{t})$ for the larger-population
activity and its marginal $\pxx(\ya_{t})$ for the sample activity at that
bin. These two distributions are related by
\begin{equation}
  \label{eq:conditional_hypergeometric_t}
  \pxx(\ya_{t})= \sum_{\yA_{t}=0}^{N} \yG_{\ya_{t}\,\yA_{t}} \;\pxxx(\yA_{t})
  \qquad \ya_{t} \in \set{0,\dotsc,n},
\end{equation}
where $\yG_{\ya\yA}$ is the hypergeometric distribution defined in
\eqn~\eqref{eq:conditional_hypergeometric}, characteristic of
\enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}.%[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.

@@

In step~\ref{item:maxent_step} we actually use the
\emph{minimum-relative-entropy} method
\citep{hobsonetal1973,csiszar1985}[\sect~5.2.2]{sivia1996_r2006} with
respect to a uniform reference distribution. We shall still call it
\enquote{maximum-entropy} for brevity. It amounts to two prescriptions:
first, take the distributions satisfying specific convex constraints --
such as fixed expectations -- and among them select the one having minimum
relative entropy with respect to a reference distribution; second, judge --
and therefore set -- those expectations to be equal to some measured
averages, typically time averages. (A judgement is required because
probability expectations and empirical averages differ numerically in
general; a misjudgement of their approximate equality can lead to
\enquote{silly answers}, to quote MacKay \citep[p.~308]{mackay1995_r2003}.)

@@

The amount and degrees of the constraining moments depend on the questions
and hypotheses under study. We give some examples in the
next two sections.

Note that the constraint equation~\eqref{eq:final_constraint} for the
distribution~\eqref{eq:app_maxent_pop_first} may not have solutions
if $N$ is strictly larger than $n$, but it's still possible to find The possible discrepancy between its
two sides typically increases with the number of constraints. This
discrepancy comes from an approximation implicit in the maximum-entropy
method combined with one assumption specific to our application: that
measured averages equal expected values (or equivalently, the number of
time bins $T$ is infinite), and that an activity level $\yA$ can equally
likely be generated by any set of $\yA$ neurons in the larger population.
The magnitude of this discrepancy can be a signature of the presence of
neuronal \enquote{assemblies} or \enquote{clusters}
\citep[\chap~12]{gerstneretal2014}{hebb1949_r2002}. We discuss this matter
in \sect~\ref{sec:assumptions}.

@@

The first assumption, of infinite exchangeability
\citep[\chap~4]{bernardoetal1994_r2000}{dawid2013} of the probability
distribution $\pf[(\yA_{1},\dotsc, \yA_{T}) \| \yI]$ for the sequence of
activities, implies the following mixture form by de~Finetti's theorem
\citep{definetti1930,hewittetal1955}:
\begin{equation}
  \label{eq:mixture_from_definetti}
%  \exp[-T\;\sh(\yF;\ynu)] \quad
%  \pxxx[(\yA_{1},\dotsc,\yA_{T})] \propto
  \pf[(\yA_{1},\dotsc, \yA_{T}) \| \yI]
  = \int\!\!\!\di\ynu\;\pf(\ynu \|\yI)\,\prod_{\yA=0}^{N}{\nu_{\yA}}^{T\,\yFF_{\yA}},
\end{equation}
where $\ynu\defd(\nu_{\yA})$ is the long-run frequency distribution of
activities in an imaginary continuation or repetition of the recording, and
$\pf(\ynu \|\yI)\;\di\ynu$ our belief distribution about it. The
integration is over an $N$-dimensional simplex.

The second assumption concerns the latter distribution. Its density
function has one of the following expressions:
\begin{enumerate}[label=(\textit{\alph*}),wide]
\item an entropic prior \citep{skillingetal1984,rodriguez1991,neumann2007}
  \begin{subequations}\label{eq:maxent_prior}
    \begin{equation}
      \label{eq:entropic_prior}
      \pf(\ynu \| \yI) \propto  \exp[-\yL\;\sh(\yF;\yr)]
      \approx \binom{L}{L\nu_{0}, \dotsc, L\nu_{N}}\;
      \prod_{\yA}{r_{\yA}}^{L\nu_{\yA}}  
    \end{equation}
    where $\yr\defd(r_{\yA})$ is a reference distribution and the
    expression in larger parentheses is a multinomial coefficient;
  \item a \enquote{sufficiency prior}, from the exponential family, which
    follows from the Pitman-Koopman theorem if some moments $\set{m}$ are
    sufficient statistics for our belief about $\ynu$:
    \begin{equation}
      \label{eq:suff_prior}
      \pf(\ynu \| \yI) \propto
      \int\!\!\!\di\yla\;\pf(\yla \| \yI)\;
      \prod_{\yA}\delt\Biggl\{\ynu_{\yA}-
      \frac{r_{\yA}}{Z(\yla)}\,\exp\Bigl[
      \tsum_{m}\lambda_{m}\tbinom{\yA}{m}/\tbinom{N}{m}
      \Bigr]\Biggr\},
    \end{equation}
  \end{subequations}
  where $\yla\defd(\lambda_{m})$ are Lagrange multipliers and $Z(\yla)$ a
  normalization factor for the exponential distribution within the delta.
  This is a formal expression, the delta symbolizing that the possible
  domain of $\ynu$ is restricted from an $N$-dimensional simplex to a
  lower-dimensional subset thereof, parametrized by $\yla$.
\end{enumerate}
